Report on the Third Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE3)

### 2015-11-05 WSSSPE3/report/wg_main_metrics.tex
small changes, line breaks

Metrics for scientific software are important for tenure and promotion, scientific impact, discovery, reducing duplication, serving as a basis for potential industrial interest in adopting software, and making a case for new or continued funding.  However, there is no commonly-used standard for collecting or presenting metrics, nor is it known if there is a common set of metrics for scientific software.  It is imperative that scientific software stakeholders understand that it is useful to collect metrics.

@@

The group discussion here focused on identifying existing frameworks and activities for scientific software metrics.  The group identified the following related activities:

@@

The group discussion began by agreeing on the common purpose of creating a set of guidance giving examples of specific metrics for the success of scientific software in use, why they were chosen, what they are useful to measure, and any challenges and pitfalls; then publish this as a white paper.  The group discussed many questions related to useful metrics for scientific software including addressing if there is a common set of metrics that can be filtered in some way, can metrics be fit into a common template, which metrics would be the most useful for each stakeholder, which metrics are the most helpful and how would we assess this, how are metrics monitored, and many more.  A more complete bulleted list of these questions can be found in Appendix H.  Next, a roadmap for how to proceed was discussed including creating a set of milestones and tasks.  The idea was put forth for the group to interact with the organizing committee of the 2016 NSF Software Infrastructure for Sustained Innovation (SI2) PI workshop in order to email out a software metrics survey to all SI2 and related awardees as a targeted and relevant set of stakeholders.  This survey would be created by one of the student group members.  Similarly, it was suggested that a software metrics survey be sent to the UK SFTF and TRDF software projects to ask them what metrics would be useful to report.  The remainder of the discussion focused mainly on the creation of a white paper on this topic.  This resulted in a paper outline and writing assignments with the goal of publishing in venues including WSSSPE4, IEEE CISE, or JORS.

>>>>

Metrics for scientific software are important for tenure and promotion, scientific impact, discovery, reducing duplication, serving as a basis for potential industrial interest in adopting software, prioritizing develop and support towards 
strategic objectives and making a case for new or continued funding.  However, there is no commonly-used 
standard for collecting or presenting metrics, nor is it known if there is a common set of metrics for scientific software. 
 It is imperative that scientific software stakeholders understand that it is useful to collect metrics.

@@

The group discussion  focused on identifying existing frameworks and activities for scientific software metrics.  
The group identified the following related activities:

@@

The group discussion began by agreeing on the common purpose of creating a set of guidance giving examples of specific metrics for the success of scientific 
software in use, why they were chosen, what they are useful to measure, and any challenges and pitfalls; then publish this as a white paper. 
 The group discussed many questions related to useful metrics for scientific software including addressing if there is a common set of metrics that 
 can be filtered in some way, can metrics be fit into a common template, which metrics would be the most useful for each stakeholder, 
 which metrics are the most helpful and how would we assess this, how are metrics monitored, and many more.  
 A more complete bulleted list of these questions can be found in Appendix H.  Next, a roadmap for how to proceed was discussed including
 creating a set of milestones and tasks.  The idea was put forth for the group to interact with the organizing committee of the 2016 NSF Software Infrastructure
  for Sustained Innovation (SI2) PI workshop in order to send a software metrics survey to all SI2 and related awardees as a targeted and relevant set of stakeholders.  
 The five solicitations for software elements released  under the NSF SI2 program all included metrics as a required component with submitters requested to include 
 {\it "a list of tangible metrics, with end user involvement, to be used to measure the success of the software element developed, ...'}. These metrics are then reported as part of annual reports to NSF by the projects. Although neither the proposal text describing the metrics nor the reported metric results are publicly available, there is reason to believe that the community will be willing to provide this information through a survey mechanism. 
This survey would be created by one of the student group members.  Similarly, it was suggested that a software metrics survey be sent to the 
UK SFTF and TRDF software projects to ask them what metrics would be useful to report.  The remainder of the discussion focused mainly on the 
creation of a white paper on this topic.  This resulted in a paper outline and writing assignments with the goal of publishing in venues including 
WSSSPE4, IEEE CISE, or JORS.
